### 2022-07-15 response-letter.md

Dear editor,

Thank you for the valuable feedback on our manuscript. We are pleased that you
and the reviewer found interest and worth in our manuscript. We addressed each comment and found that the manuscript is indeed clearer now, and we hope to have it approved after your consideration.

We have added more context in the introduction section, but we've found it hard
to expand the first paragraph. We believe that this particular paragraph needs
to be more succinct, at the same time that it provides a complete picture of the
paper, framed in a storytelling structure. Then, on the second paragraph, we
start providing literature background and many examples, to help the reader
understand where our paper stands in the scientific context. We have included
discussions about the quality and accuracy of range maps (lines XX-XX), examples
of data biases (lines XX-XX), and we addressed the uniqueness of our study
compared to previous research (lines XX-XX). We hope that this structure is
clarifying and appealing to the readers.

Regarding the GBIF data cleaning, retrieval and the validation steps, we added a
more detailed description of our process (lines XX-XX, XX-XX and XX-XX). Data
retrieving is detailed in the source code provided on our GitHub repository
(code/09-GBIF_data.jl). We used the list of species from the network dataset and
retrieved any occurrence within our bounding box (between longitudes -20.0 and
55.0 degrees; and latitudes -35.0 and 40.0 degrees). We changed a two things
regarding geographical information during data retrieval:
1. We had originally added an argument to capture only the data that had
   information about the continent location equals to "AFRICA". However, this
   constrained our search by thousands of occurrences, and added a filtering
   layer that could be interpreted as "data cleaning". We removed that because
   we don't perform any kind of data cleaning with the IUCN dataset.

2. We added, however, a constraint to exclude occurrences falling within a 4°x4°
   degrees square around the "Null Island" (see code/09-GBIF_data.jl#L42 on our
   repository). Rather than a data cleaning step, we see this as a geographic
   filter complementing the bounding box we used to retrieve data from GBIF and
   to limit IUCN range maps. This is important because the existence of
   coordinates is a fundamental part of our analysis, and a (00,00) coordinate
   can be interpreted as a null coordinate.

Some taxonomic investigation needed to be done due to inconsistencies between
GBIF and IUCN nomenclatures which would block our analysis completely (in the case of _Taurotragus oryx_). However, in the case of *Canis aureus*, as discussed in the paper, we could not track wich occurrences or which maps corresponded to which species, which we think is a great illustration of our point. We didn't
perform any further data cleaning procedures because this could hinder the goals
of the paper, which was to investigate the characteristics of the data available
for research. As we needed to compare two sets of raw data, and data cleaning
protocols are dependent on the use objectives, we chose to avoid data
manipulation both for GBIF and IUCN datasets.

@@

For the validation step, we understand our phrasing was confusing, and we have
rewritten this section to clarify our method. Rather than using
absence data, what we did was to transform the point data into raster files, and
then restrict our range maps to the locations where we had GBIF occurrences.
Therefore, the pixels where there are no GBIF occurrences are not considered
true absence of a species, but absence of **data** about the occurrence of that
species.

Finally, we have addressed all the other minor comments throughout the text as
described below:
- We've added citations in the Methods section to support our argument that a
  mismatch between ranges can be originated from different sources (L. 68-70)
- We have suppressed lines 214-217.
- The abstract is now restructured according to PeerJ guidelines.
- References to tables and figures in the Discussion section were suppressed,
  with the exception of Figure 05 (line XX), which is an example of our results
  to clarify an important point of the discussion.
- Lines 22-23 (now lines XX-XX): the cited literature are not examples of models
  that do not take ecological interactions into account, but studies that have
  demonstrated how ecological interactions might be responsible for shaping a
  species' distribution. We've rewritten the sentence to clarify that.
-  Lines 68-70: we have rewritten the sentence to include the consideration
   misestimation of both species in a trophic interaction. We have also added
   citations and examples.
- We have addressed the grammatical suggestions throughout the text, except for
  the legend of figure 1. We thought that the suggested phrasing adds a little
  bit of confusion to the comprehension of the figure.

### 2022-07-15 draft.tex
Deploying to gh-pages from @ graciellehigino/ms_range_interactions@f0df5760733f9e4661a0419c313416a90f10cd3f 🚀

Interactions form complex networks that shape ecological structures and
maintain the essential functions of ecosystems, such as seed dispersal,
pollination, and biological control (Albrecht 2018; Fricke et al. 2022)
that ultimately affect the composition, richness, and successional
patterns of communities across biomes. Yet, the connection between
occurrence and interaction data is a frequent debate in ecology
{[}Blanchet, Cazelles, and Gravel (2020); Wisz2013RolBio{]}. For
instance, macroecological models are often used with point or range
occurrence data in order to investigate the dynamics of a species with
its environment. However, these models do not account for ecological
interactions, although it has been demonstrated that they might largely
affect species distribution {[}Abrego et al. (2021); Afkhami, McIntyre,
and Strauss (2014); Araújo, Marcondes-Machado, and Costa (2014); Godsoe
et al. (2017); Godsoe and Harmon (2012); Gotelli, Graves, and Rahbek
(2010); Wisz2013RolBio{]}. Some researchers argue that occurrence data
can also capture real-time interactions (see Roy et al. 2016; Ryan et
al. 2018), and, because of that, it would not be necessary to include
ecological interaction dynamics in macroecological models. On the other
hand, many mechanistic simulation models in ecology have considered the
effect of competition and facilitation in range shifts. For example,
Gotelli \emph{et al.} -(Gotelli, Graves, and Rahbek 2010) demonstrate
how conspecific attraction might be the main factor driving the
distribution of migratory birds; Afkhami \emph{et al.} -(Afkhami,
McIntyre, and Strauss 2014) explores how mutualistic fungal endophytes
are responsible or expanding the range of native grass; many other
examples are discussed in Wisz \emph{et al.} -(Wisz et al. 2013).
Although interactions across trophic levels are demonstrated to
determine species range (Wisz et al. 2013), the use of these
interactions in mechanistic simulation models in macroecology remains
insufficient (as discussed in Cabral, Valente, and Hartig 2017).

A significant challenge in this debate is the quality and quantity of
species distribution and ecological data (Boakes et al. 2010; Ronquillo
et al. 2020; Meyer, Weigelt, and Kreft 2016) - a gap that can lead to
erroneous conclusions in macroecological research (Hortal et al. 2008).
Amongst the geographical data available are the range maps provided by
the International Union for the Conservation of Nature (IUCN). Such maps
consist of simplified polygons, often created as alpha or convex hulls
around known species locations, refined by expert knowledge about the
species (IUCN Red List Technical Working Group 2019). These maps can be
used in macroecological inferences in the lack of more precise
information (Fourcade 2016; Alhajeri and Fourcade 2019), but it has been
recommended that they are used with caution since they tend to
underestimate the distribution of species that are not well-known
(Herkt, Skidmore, and Fahr 2017) (especially at fine scale resolutions;
(Hurlbert and Jetz 2007; Hurlbert and White 2005)), do not represent
spatial variation in species occurrence and abundance (Dallas, Pironon,
and Santini 2020), and can include inadequate areas within the estimated
range. Another source of species distribution information is the Global
Biodiversity Information Facility (GBIF), which is an online repository
of georeferenced observational records that come from various sources,
including community science programs, museum collections, and long-term
monitoring schemes. A great source of bias in these datasets is the
irregular sampling effort, with more occurrences originating from
attractive and accessible areas and observation of charismatic species
(Alhajeri and Fourcade 2019). As for ecological data, a complete
assessment is difficult and is aggravated by biased sampling methods and
data aggregation (Poisot et al. 2020; Hortal et al. 2015). Nevertheless,
we have witnessed an increase in the availability of biodiversity data
in the last decades, including those collected through community science
projects (Callaghan et al. 2019; Pocock et al. 2015) and dedicated
databases, such as Mangal (Poisot et al. 2016). This provides an
opportunity to merge species distribution and ecological interaction
data to improve our predictions of where a species may be found across
large spatial scales (e.g., continental and global).

@@

Organisms cannot persist unless they are directly or indirectly
connected to a primary producer within their associated food web (Power
1992). Therefore, the range of a predator (omnivore or carnivore)
depends on the overlapping ranges of its preys. If sections of a
predator's range does not overlap with at least one of its prey it will
become disconnected from primary producers, and therefore we would not
expect the predator to occur in this area. This mismatch can be the
result of different mechanisms, like the misestimation of both the
predator's and the preys' ranges {[}Ladle and Hortal (2013);
Rondinini2006TraDif{]}, taxonomic errors (Isaac, Mallet, and Mace 2004;
Ladle and Hortal 2013), or the lack of information about trophic links
(i.e., the lack of connection between the ranges of a predator and a
primary producer may be due a third species we don't know is connected
to both). Thus, given that herbivores are the main connection between
plant resources (directly limited by environmental conditions) and
predators (Dobson 2009; Scott et al. 2018), here we adjusted the ranges
of predators based on a simple rule: we removed any part of a predator's
range that did not intersect with the range of at least one prey
herbivore species. So, unless the range of the predator overlapped with
at least one prey item, which in turn is directly connected to a primary
producer (plants), we removed that section of the predator's range.
Finally, we calculated the difference in range size between the original
IUCN ranges and those adjusted based on species interaction data.

@@

The proportion of GBIF pixels (pixels with at least one GBIF occurrence)
falling in the IUCN ranges varied from low to high depending on the
species (fig.~\ref{fig:gbif}, left). The lowest proportions occurred for
species with small ranges (such as \emph{Lycaon pictus}), although some
species with small ranges showed high overlap. Species with median and
large ranges had high proportions of occurrences falling into their IUCN
range. Predators and preys displayed similar overlap variations. While
no species had all of its GBIF occurrences within its IUCN range, one
species had this proportion equal to zero, \emph{Canis aureus}, which is
also the only species whose range is not covered by any of its preys.
This result reinforces the concern raised in the literature on the use
of IUCN range maps for species that are not well known (Herkt, Skidmore,
and Fahr 2017), demonstrating how small range species are likely to have
their distribution underestimated in the IUCN database. Additionally,
the fact that \emph{Canis aureus} had none of its GBIF pixels
overlapping with IUCN maps suggests a taxonomic mismatch between both
databases, which we explore in the Discussion section.

The proportion of GBIF pixels in updated ranges can only be equal to or
lower than that of the original ranges, as our analysis removes pixels
from the original range and does not add new ones. Rather, the absence
of a difference between the two types of ranges indicates that no pixels
with GBIF observations, hence likely true habitats, were removed by our
analysis. Here this proportion was mostly similar to that of the
original IUCN ranges for most predator species (fig.~\ref{fig:gbif}).
Four species showed no difference in proportion while three species
showed only small differences (proportions of 0.01 to 0.05). On the
other hand, two species, \emph{Canis mesomelas}, and \emph{Leptailurus
serval} showed very high differences, with overlaps lower by 0.548 and
0.871 respectively. For \emph{Leptailurus serval}, none of the GBIF
observations occurred in the updated range. These two species are also
the only predators with a single prey in our metaweb. Our results
delineate how a mismatch between GBIF and IUCN databases differ greatly
with small changes in herbivore species ranges, and it is somewhat
positively related to range size for predator species. Moreover, we show
that accounting for interactions does not necessarily aggravates this
dissimilarity, but it is relevant for species with little ecological
information or specialists.

@@

The jackal is a widespread taxon in northern Africa, Europe, and
Australasia, generally well adapted to local conditions due to its
largely varied diet (Tsunoda and Saito 2020; Krofel et al. 2021).
Because of that, we expected that the \emph{Canis} species in our
dataset would be the ones losing the least amount of range, with a
higher value of the proportion of GBIF pixels within their IUCN range
maps. However, the taxonomy of this group is a matter of intense
discussion, as molecular and morphological data seem to disagree in the
clustering of species and subspecies (Krofel et al. 2021; Stoyanov
2020). This debate is indeed reflected in our analysis: the GBIF
identification of the golden jackal is incompatible with the one used by
IUCN, each of them mapping its distribution in completely different
places. This led to a complete exclusion of \emph{Canis aureus} from its
original range in our analysis, despite the fact that this species has
four documented preys in our metaweb. This example illustrates how the
taxonomic, geographical and ecological data can be used to validate one
another.

@@

In the Serengeti food web there is a positive relationship between the
out degrees of predators and the size of their ranges. In addition, our
results show that there is a negative relationship between the relative
loss of predators' ranges and their number of preys, reinforcing the
idea that generalist species can preserve their distributions longer
while losing interactions. The factors limiting the geographical range
of a species in a community can vary with connectivity and richness
(Svenning et al. 2014). Younger communities may be more affected by
environmental limitations because they are dominated by generalist
species, while older metacommunities are probably affected in different
ways in the center of the distribution, at the edge of ranges, and in
sink and source communities (Svenning et al. 2014; Godsoe et al. 2017;
Cazelles et al. 2016; Bullock et al. 2000). Additionally, it is likely
that species with larger ranges of distribution and those that are more
generalists would co-occur with a greater number of other species
(Dáttilo et al. 2020), while dispersal capacity of competitive species
modulate their aggregation in space and the effect of interactions on
their range limits (Godsoe et al. 2017).

@@

The lack of superposition between IUCN range maps and GBIF occurrences
suggests that we certainly do miss geographical information about the
distribution of a certain species, but it is not an indicator of the
completeness of the information about ecological interactions. However,
if both GBIF and IUCN occurrences tend to superpose and still the
species is locally removed, this indicates we don't have information
about all its interactions. The combination of this rationale with our
method of updating range maps based on ecological interactions allows us
to have a clearer idea of which information we are missing. For example,
the lion (\emph{Panthera leo}) was one of the species with the smallest
difference between the original and the updated ranges, but 59.5\% of
the GBIF occurrences for this species fell outside the IUCN range
(fig.~\ref{fig:gbif}). In this particular case, the IUCN maps seem to
agree with species interaction data. However, the disagreement between
the IUCN and the GBIF databases is concerning and suggests that the IUCN
maps might underestimate the lion's distribution. On the other hand,
\emph{Leptailurus serval} and \emph{Canis mesomelas} are two of the
three species that lose the higher proportion of range due to the lack
of paths to a herbivore, but are also some of the species with the
higher proportion of GBIF occurrences inside IUCN range maps
(fig.~\ref{fig:gbif}). This indicates that the information we are
missing for these two species is related to either the occurrence of an
interaction or the presence of interacting species. To illustrate that,
we mapped the GBIF data for the prey of \emph{Leptailurus serval}, with
a mobility buffer around each point (fig.~\ref{fig:serval}). When
considering GBIF data, approximately 53\% of the prey's occurrences are
within the portion of the serval's range that was lost. With the buffer
area, this corresponds to 13\% of the lost range. This means that by
adding GBIF information, we would reduce the loss of range (or
information) for the predator by 13\% since its distribution is
conditional on the occurrence of its preys.

@@

Finally, the extreme case of \emph{Canis aureus} illustrates a lack of
both geographical and ecological information: none of its GBIF
occurrences and none of its preys occur inside its IUCN range. We
believe, therefore, that the validation of species distribution based on
ecological interaction is a relevant method that can further fill in
information gaps. Nevertheless, it is imperative that more
geographically explicit data about ecological networks and interactions
become available. This would help clarify when cooccurrences can be
translated into interactions and help the development of more advanced
validation methods for occurrence data.

@@

Here we demonstrated how we can detect uncertainty in species
distribution data using ecological interactions. Knowing where
questionable occurrence data are can be crucial in ecological modelling
(Hortal 2008; Ladle and Hortal 2013), and accounting for these errors
can improve model outputs by diminishing the error propagation (Draper
1995). For instance, we believe this is a way to account for ecological
interactions in habitat suitability models without making the models
more complex, but by making sure (not assuming) that the input data -
the species occurrence - actually accounts for ecological interactions.
It is important to notice, however, that the quality and usefulness of
this method are highly correlated with the amount and quality of data
available about species' occurrences \textbf{and} interactions. With
this paper, we hope to add to the collective effort to decode the
encrypted message that is the occurrence of a species in space and time.
A promising avenue that adds to our method is the prediction of networks
and interactions in large scales (Strydom et al. 2021), for they can add
valuable information about ecological interactions where they are
missing. Additionally, in order to achieve a robust modelling framework
towards actual species distribution models we should invest in efforts
to collect and combine open data on species occurrence and interactions,
especially because we may be losing ecological interactions at least as
fast as we are losing species (Valiente-Banuet et al. 2015).

@@

We acknowledge that this study was conducted on land within the
traditional unceded territory of the Saint Lawrence Iroquoian,
Anishinabewaki, Mohawk, Huron-Wendat, and Omàmiwininiwak nations. GH,
FB, GD, and NF are funded by the NSERC Computational Biodiversity
Science and Services (BIOS\(^2\)) CREATE program; FB, NF, and TP are
funded by the Institute for Data Valorization (IVADO); NF and TP are
funded by a donation from the Courtois Foundation; GD is funded by the
FRQNT doctoral scholarship; TP is funded by the Canadian Institute of
Ecology \& Evolution; FW is funded by the Royal Society (Grant number:
CHL\textbackslash R1\textbackslash180156).

@@

@@

>>>>

Interactions form complex networks that shape ecological structures and
maintain the essential functions of ecosystems, such as seed dispersal,
pollination, and biological control (Albrecht 2018; Fricke et al. 2022)
that ultimately affect the composition, richness, and successional
patterns of communities across biomes. Yet, the connection between
occurrence and interaction data is a frequent debate in ecology
(Blanchet, Cazelles, and Gravel 2020; Wisz et al. 2013). For instance,
macroecological models are often used with point or range occurrence
data in order to investigate the dynamics of a species with its
environment. However, these models do not account for ecological
interactions, although it has been demonstrated that they might largely
affect species distribution (Abrego et al. 2021; Afkhami, McIntyre, and
Strauss 2014; Araújo, Marcondes-Machado, and Costa 2014; Godsoe et al.
2017; Godsoe and Harmon 2012; Gotelli, Graves, and Rahbek 2010; Wisz et
al. 2013). Some researchers argue that occurrence data can also capture
real-time interactions (see Roy et al. 2016; Ryan et al. 2018), and,
because of that, it would not be necessary to include ecological
interaction dynamics in macroecological models. On the other hand, many
mechanistic simulation models in ecology have considered the effect of
competition and facilitation in range shifts. For example, Gotelli
\emph{et al.} (2010) demonstrate how conspecific attraction might be the
main factor driving the distribution of migratory birds; Afkhami
\emph{et al.} (2014) explores how mutualistic fungal endophytes are
responsible or expanding the range of native grass; many other examples
are discussed in Wisz \emph{et al.} (2013). Although interactions across
trophic levels are demonstrated to determine species range (Wisz et al.
2013), the use of these interactions in mechanistic simulation models in
macroecology remains insufficient (as discussed in Cabral, Valente, and
Hartig 2017).

A significant challenge in this debate is the quality and quantity of
species distribution and ecological data (Boakes et al. 2010; Ronquillo
et al. 2020; Meyer, Weigelt, and Kreft 2016) - a gap that can lead to
erroneous conclusions in macroecological research (Hortal et al. 2008).
Amongst the geographical data available are the range maps provided by
the International Union for the Conservation of Nature (IUCN). Such maps
consist of simplified polygons, often created as alpha or convex hulls
around known species locations, refined by expert knowledge about the
species (IUCN Red List Technical Working Group 2019). These maps can be
used in macroecological inferences in the lack of more precise
information (Fourcade 2016; Alhajeri and Fourcade 2019), but it has been
recommended that they are used with caution since they tend to
underestimate the distribution of species that are not well-known
(Herkt, Skidmore, and Fahr 2017) (especially at fine scale resolutions;
Hurlbert and Jetz (2007); Hurlbert and White (2005)), do not represent
spatial variation in species occurrence and abundance (Dallas, Pironon,
and Santini 2020), and can include inadequate areas within the estimated
range. Another source of species distribution information is the Global
Biodiversity Information Facility (GBIF), which is an online repository
of georeferenced observational records that come from various sources,
including community science programs, museum collections, and long-term
monitoring schemes. A great source of bias in these datasets is the
irregular sampling effort, with more occurrences originating from
attractive and accessible areas and observation of charismatic species
(Alhajeri and Fourcade 2019). As for ecological data, a complete
assessment is difficult and is aggravated by biased sampling methods and
data aggregation (Poisot et al. 2020; Hortal et al. 2015). Nevertheless,
we have witnessed an increase in the availability of biodiversity data
in the last decades, including those collected through community science
projects (Callaghan et al. 2019; Pocock et al. 2015) and dedicated
databases, such as Mangal (Poisot et al. 2016). This provides an
opportunity to merge species distribution and ecological interaction
data to improve our predictions of where a species may be found across
large spatial scales (e.g., continental and global).

@@

Organisms cannot persist unless they are directly or indirectly
connected to a primary producer within their associated food web (Power
1992). Therefore, the range of a predator (omnivore or carnivore)
depends on the overlapping ranges of its preys. If sections of a
predator's range does not overlap with at least one of its prey it will
become disconnected from primary producers, and therefore we would not
expect the predator to occur in this area. This mismatch can be the
result of different mechanisms, like the misestimation of both the
predator's and the preys' ranges (Ladle and Hortal 2013; Rondinini et
al. 2006), taxonomic errors (Isaac, Mallet, and Mace 2004; Ladle and
Hortal 2013), or the lack of information about trophic links (i.e., the
lack of connection between the ranges of a predator and a primary
producer may be due a third species we don't know is connected to both).
Thus, given that herbivores are the main connection between plant
resources (directly limited by environmental conditions) and predators
(Dobson 2009; Scott et al. 2018), here we adjusted the ranges of
predators based on a simple rule: we removed any part of a predator's
range that did not intersect with the range of at least one prey
herbivore species. So, unless the range of the predator overlapped with
at least one prey item, which in turn is directly connected to a primary
producer (plants), we removed that section of the predator's range.
Finally, we calculated the difference in range size between the original
IUCN ranges and those adjusted based on species interaction data.

@@

The proportion of GBIF pixels (pixels with at least one GBIF occurrence)
falling within the IUCN ranges varied from low to high depending on the
species (fig.~\ref{fig:gbif}, left). The lowest proportions occurred for
species with small ranges. Amongst herbivores, \emph{Rhabdomys pumilio}
has a proportion of 22.6\% of its presence pixels within its IUCN range,
while predators have this proportion above 55\% (such as \emph{Lycaon
pictus}, with 55.1\%, and \emph{Canis aureus}, with 56.2\%).
Nevertheless, some species with smaller ranges showed high data overlap
(such as \emph{Canis mesomelas}, with 94.9\%, and many herbivores).
Overall, predators and preys displayed similar overlap variations, and
species with median and large ranges had higher proportions of
occurrences falling into their IUCN range.

The proportion of GBIF pixels in updated ranges can only be equal to or
lower than that of the original ranges, as our analysis removes pixels
from the original range and does not add new ones. Rather, the absence
of a difference between the two types of ranges indicates that no pixels
with GBIF observations, hence likely true habitats, were removed by our
analysis. Here this proportion was mostly similar to that of the
original IUCN ranges for most predator species (fig.~\ref{fig:gbif}).
Two species showed no difference in proportion while four species showed
only small differences (\emph{Crocuta crocuta} lost 1.3\% of the
original data overlap; \emph{Acinonyx jubatus} lost 1.9\%;
\emph{Panthera pardus} lost 8.8\%; and \emph{Caracal caracal} lost
12.3\%). \textless!---\textgreater Note to self: here I calculated how
much the difference represents as a proportion of the total proportion,
which means the decreased rate. E.g., when the range\_prop\_diff is
equal to the original proportion (range\_prop), this means that this
species lost 100\% of the original proportion. Therefore I can't use
range\_prop\_diff numbers if I'm talking about \%
lost.\textless---\textgreater{} On the other hand, three species,
\emph{Canis aureus}, \emph{Canis mesomelas}, and \emph{Leptailurus
serval} showed very high differences, with overlaps lowered by 100\%,
57.4\%, and 100\% respectively. These last two species are also the only
predators with a single prey in our metaweb. \emph{Canis aureus} has
four preys, but it has one of the smallest ranges in IUCN, which is not
covered by any of its preys. This result reinforces the concern raised
in the literature on the use of IUCN range maps for species that are not
well known (Herkt, Skidmore, and Fahr 2017), demonstrating how small
range species are likely to have their distribution underestimated in
the IUCN database. Additionally, the fact that \emph{Canis aureus} had
such a conspicuous mismatch between both the original and updated IUCN
range maps, and between GBIF and IUCN data, may indicate a taxonomic
incongruency between the three databases used here, which we explore in
the Discussion section. Our results delineate how a mismatch between
GBIF and IUCN databases differ greatly with small changes in herbivore
species ranges, and it is somewhat positively related to range size for
predator species. Moreover, we show that accounting for interactions
does not necessarily aggravates this dissimilarity, but it is relevant
for species with little ecological information or specialists.

@@

The jackal is a widespread taxon in northern Africa, Europe, and
Australasia, generally well adapted to local conditions due to its
largely varied diet (Tsunoda and Saito 2020; Krofel et al. 2021).
Because of that, we expected that the \emph{Canis} species in our
dataset would be the ones losing the least amount of range, with a
higher value of the proportion of GBIF pixels within their IUCN range
maps. However, the taxonomy of this group is a matter of intense
discussion, as molecular and morphological data seem to disagree in the
clustering of species and subspecies (Krofel et al. 2021; Stoyanov
2020). This debate probably influenced our results: with originally only
56.2\% of the GBIF pixels of the golden jackal (\emph{Canis aureus})
overlapping with the IUCN data, we suspect that many of the GBIF
occurrences refer to other \emph{Canis} species, and that its taxonomic
identification in the network database is probably outdated. This led to
a complete exclusion of \emph{Canis aureus} from its original range in
our analysis, despite the fact that this species has four documented
preys in our metaweb. This example illustrates how the taxonomic,
geographical and ecological data can be used to validate one another.

@@

In the Serengeti food web there is a positive relationship between the
out degrees of predators and the size of their ranges. Here, we showed
that there is a negative relationship between the relative loss of
predators' ranges and their number of preys, reinforcing the idea that
generalist species can preserve their distributions longer while losing
interactions. The factors limiting the geographical range of a species
in a community can vary with connectivity and richness (Svenning et al.
2014). Younger communities may be more affected by environmental
limitations because they are dominated by generalist species, while
older metacommunities are probably affected in different ways in the
center of the distribution, at the edge of ranges, and in sink and
source communities (Svenning et al. 2014; Godsoe et al. 2017; Cazelles
et al. 2016; Bullock et al. 2000). Additionally, it is likely that
species with larger ranges of distribution and those that are more
generalists would co-occur with a greater number of other species
(Dáttilo et al. 2020), while dispersal capacity of competitive species
modulate their aggregation in space and the effect of interactions on
their range limits (Godsoe et al. 2017).

@@

The lack of superposition between IUCN range maps and GBIF occurrences
suggests that we certainly do miss geographical information about the
distribution of a certain species, but this is not an indicator of the
completeness of the information about ecological interactions. However,
if both GBIF and IUCN occurrences tend to superpose and still the
species is locally removed, this indicates we don't have information
about all its interactions. The combination of this rationale with our
method of updating range maps based on ecological interactions allows us
to have a clearer idea of which information we are missing. For example,
the lion (\emph{Panthera leo}) was one of the species with no difference
between the original and the updated ranges, but 40.7\% of the GBIF
occurrences for this species fell outside its IUCN range
(fig.~\ref{fig:gbif}). In this particular case, the IUCN maps seem to
agree with species interaction data. However, the disagreement between
the IUCN and the GBIF databases is concerning and suggests that the IUCN
maps might underestimate the lion's distribution. On the other hand,
\emph{Leptailurus serval} and \emph{Canis mesomelas} are two of the
three species that lose the higher proportion of range due to the lack
of paths to a herbivore, but are also some of the species with the
higher proportion of GBIF occurrences inside IUCN range maps
(fig.~\ref{fig:gbif}). This indicates that the information we are
missing for these two species is related to either the occurrence of an
interaction or the presence of interacting species. To illustrate that,
we mapped the GBIF data for the prey of \emph{Leptailurus serval}, with
a mobility buffer around each point (fig.~\ref{fig:serval}). When
considering GBIF data, approximately 42\% of the prey's occurrences are
within the portion of the serval's range that was lost. With the buffer
area, this corresponds to 15\% of the lost range. This means that by
adding GBIF information, we would reduce the loss of range (or
information) for the predator by 15\% since its distribution is
conditional on the occurrence of its preys.

@@

Finally, the extreme case of \emph{Canis aureus} illustrates a lack of
both geographical and ecological information: only half of its GBIF
presence pixels and none of its preys occur inside its IUCN range. We
believe, therefore, that the validation of species distribution based on
ecological interaction is a relevant method that can further fill in
information gaps. Nevertheless, it is imperative that more
geographically explicit data about ecological networks and interactions
become available. This would help clarify when cooccurrences can be
translated into interactions (Windsor et al. 2022) and help the
development of more advanced validation methods for occurrence data.

@@

Here we demonstrated how we can detect uncertainty in species
distribution data using ecological interactions. Knowing where
questionable occurrence data are can be crucial in ecological modelling
(Hortal 2008; Ladle and Hortal 2013), and accounting for these errors
can improve model outputs by diminishing the error propagation (Draper
1995). For instance, we believe this is a way to account for ecological
interactions in habitat suitability models without making the models
more complex, but by making sure (not assuming) that the input data -
the species occurrence - actually accounts for ecological interactions.
It is important to notice, however, that the quality and usefulness of
this method are highly correlated with the amount and quality of data
available about species' occurrences \textbf{and} interactions. With
this paper, we hope to add to the collective effort to decode the
encrypted message that is the occurrence of a species in space and time.
A promising avenue that adds to our method is the prediction of networks
and interactions in large scales (Strydom et al. 2021; Windsor et al.
2022), for they can add valuable information about ecological
interactions where they are missing. Additionally, in order to achieve a
robust modelling framework towards actual species distribution models we
should invest in efforts to collect and combine open data on species
occurrence and interactions (Windsor et al. 2022), especially because we
may be losing ecological interactions at least as fast as we are losing
species (Valiente-Banuet et al. 2015).

@@

We acknowledge that this study was conducted on land within the
traditional unceded territory of the Saint Lawrence Iroquoian,
Anishinabewaki, Mohawk, Huron-Wendat, and Omàmiwininiwak nations. GH,
FB, GD, and NF are funded by the NSERC Computational Biodiversity
Science and Services (BIOS\(^2\)) CREATE program; FB, NF, and TP are
funded by the Institute for Data Valorization (IVADO); NF and TP are
funded by a donation from the Courtois Foundation; GD is funded by the
Natural Sciences and Engineering Research Council (NSERC) and the Fond
de recherche du Québec - Nature et technologie (FRQNT) doctoral
scholarships; TP is funded by the Canadian Institute of Ecology \&
Evolution; FW is funded by the Royal Society (Grant number:
CHL\textbackslash R1\textbackslash180156). We thank the editor and
reviewer for their thoughtful comments, which considerably improved this
manuscript.

@@

\leavevmode\hypertarget{ref-Rondinini2006TraDif}{}%
Rondinini, Carlo, Kerrie A. Wilson, Luigi Boitani, Hedley Grantham, and
Hugh P. Possingham. 2006. {``Tradeoffs of Different Types of Species
Occurrence Data for Use in Systematic Conservation Planning.''}
\emph{Ecology Letters} 9 (10): 1136--45.
\url{https://doi.org/10.1111/j.1461-0248.2006.00970.x}.

@@

\leavevmode\hypertarget{ref-Windsor2022UsiEco}{}%
Windsor, Fredric M., Johan van den Hoogen, Thomas W. Crowther, and
Darren M. Evans. 2022. {``Using Ecological Networks to Answer Questions
in Global Biogeography and Ecology.''} \emph{Journal of Biogeography}
n/a (n/a). \url{https://doi.org/10.1111/jbi.14447}.
